Enforced Invariant Confluence

### 2018-05-01 doc/paper/sections/evaluation.tex
took a pass

We can use 2P-sets to model a simple relational database with foreign key
constraints. Let object $O = (X, Y) = ((A_X, R_X), (A_Y, R_Y))$ consist of a
pair of two 2P-Sets $X$ and $Y$, which we view as relations. Our invariant $X
\subseteq Y$ (i.e. $(A_X - R_X) \subseteq (A_Y - R_Y)$) models a foreign key
constraint from $X$ to $Y$. We ran our decision procedure on the object with
initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and with
transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
After less than a tenth of a second, the decision procedure produced a
reachable counterexample witnessing the fact that the object is not
\invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$
can lead to a state which violates the invariant. This object is not
\invariantconfluent{} and therefore not \invariantclosed{}. Thus, previous
tools depending on \invariantclosure{} as a sufficient condition would be
unable to conclude definitively that the object is \emph{not}
\invariantconfluent{}.

@@

We then segmented our object as follows. The first segment $(\setst{(B, w)}{w =
\bot}, \setst{t_b}{b \in \ints})$ allows bidding so long as the bid is open.
The second segment $(\setst{B, w}{w \neq \bot} \cap I, \emptyset)$ includes all
auctions which have already been closed and forbids all transactions. This
segmentation captures the intuition that bids should be permitted only when the
auction is open. We ran our segmented \invariantconfluence{} decision procedure
on this example, and it was able to deduce without any human interaction that
the example was segmented \invariantconfluent{} in less than a tenth of a
second.

@@

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
\invariantconfluence{} as compared to the performance of replicating it with
eventual consistency or linearizability.
%
There are two hypotheses about the performance of segmented
\invariantconfluent{} replication that we aim to confirm.
%
First, segmented \invariantconfluent{} replication provides higher throughput
and better scalability than linearizabile replication for workloads which
require little coordination (i.e.\ low-coordination workloads).
%
Second, the throughput and scalability of segmented \invariantconfluent{}
replication decreases as we increase the fraction of transactions which require
coordination.

These hypotheses state that segmented \invariantconfluent{} is more performant
than linearizabile replication for low-coordination workloads. But by how much?
We also aim to measure the absolute performance and scalability benefits of
segmented \invariantconfluent{} replication and how they vary as we vary the
coordination required by a workload. We perform two controlled microbenchmarks
to confirm our hypotheses and discover the absolute performance benefits. The
workloads themselves are trivial, but also not salient. Our objective is to
obtain a controlled measure of throughput and scalability as we vary workload
contention.

@@

  These results offer two insights.
  %
  First, the relationship between segmented \invariantconfluent{} and
  linearizable replication is analogous to the relationship between pessimistic
  and optimistic concurrency control protocols with coordination playing the
  role of contention. Linearizable replication pessimistically assumes that
  concurrently executing \emph{any} pair of transactions will lead to an
  invariant violation. Thus, clients send transactions directly to a leader to
  be linearized. Conversely, segmented \invariantconfluent{} replication
  optimistically attempts to perform every transaction locally and without
  coordination. A server only initiates a round of coordination if it is found
  to be necessary. As a consequence, segmented \invariantconfluent{}
  replication can offer substantial performance benefits over linearizable
  replication for low-coordination workloads, but is inferior for medium to
  high contention workloads.

@@

\begin{benchmark}
  In this benchmark, we measure the scale-out of segmented
  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} while
  we vary the number of servers that we use to replicate our object. When we
  replicate with $n$ servers, we use $3n$ clients (the $3$ colocated clients on
  each server) as part of the workload. The results are shown in
  \figref{VaryNodes}.

>>>>

We can use 2P-sets to model a simple relational database with foreign key
constraints. Let object $O = (X, Y) = ((A_X, R_X), (A_Y, R_Y))$ consist of a
pair of two 2P-Sets $X$ and $Y$, which we view as relations. Our invariant $X
\subseteq Y$ (i.e. $(A_X - R_X) \subseteq (A_Y - R_Y)$) models a foreign key
constraint from $X$ to $Y$. We ran our decision procedure on the object with
initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and with
transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
After less than a tenth of a second, the decision procedure produced a
reachable counterexample witnessing the fact that the object is not
\invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$
can lead to a state that violates the invariant. This object is not
\invariantconfluent{} and therefore not \invariantclosed{}. Thus, previous
tools depending on \invariantclosure{} as a sufficient condition would be
unable to conclude definitively that the object is \emph{not}
\invariantconfluent{}.

@@

We then segmented our object as follows. The first segment $(\setst{(B, w)}{w =
\bot}, \setst{t_b}{b \in \ints})$ allows bidding so long as the bid is open.
The second segment $(\setst{B, w}{w \neq \bot} \cap I, \emptyset)$ includes all
auctions that have already been closed and forbids all transactions. This
segmentation captures the intuition that bids should be permitted only when the
auction is open. We ran our segmented \invariantconfluence{} decision procedure
on this example, and it was able to deduce without any human interaction that
the example was segmented \invariantconfluent{} in less than a tenth of a
second.

@@

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
\invariantconfluence{} as compared to the performance of replicating it with
eventual consistency or linearizability.
%
There are two hypotheses about the performance of segmented
\invariantconfluent{} replication that we aim to confirm.
%
First, segmented \invariantconfluent{} replication provides higher throughput
and better scalability than linearizable replication for workloads that
require little coordination (i.e.\ low-coordination workloads).
%
Second, the throughput and scalability of segmented \invariantconfluent{}
replication decreases as we increase the fraction of transactions that require
coordination.

These hypotheses state that segmented \invariantconfluent{} is more performant
than linearizable replication for low-coordination workloads. But by how much?
We also aim to measure the absolute performance and scalability benefits of
segmented \invariantconfluent{} replication and how they vary as we vary the
coordination required by a workload. We perform two controlled microbenchmarks
to confirm our hypotheses and discover the absolute performance benefits. The
workloads themselves are trivial, but also not salient. \jmh{Salient to what? I'm confused by the last sentence.} Our objective is to
obtain a controlled measure of throughput and scalability as we vary workload
contention.

@@

  These results offer two insights.
  %
  First, the relationship between segmented \invariantconfluent{} and
  linearizable replication is analogous to the relationship between optimistic
  and pessimistic concurrency control protocols with coordination playing the
  role of contention \jmh{I didn't understand that last clause about ``coordination playing the role of contention.} Linearizable replication pessimistically assumes that
  concurrently executing \emph{any} pair of transactions will lead to an
  invariant violation. Thus, clients send transactions directly to a leader to
  be linearized. Conversely, segmented \invariantconfluent{} replication
  optimistically attempts to perform every transaction locally and without
  coordination. A server only initiates a round of coordination if it is found
  to be necessary. As a consequence, segmented \invariantconfluent{}
  replication can offer substantial performance benefits over linearizable
  replication for low-coordination workloads, but is inferior for medium to
  high contention workloads. \jmh{say why}

@@

\begin{benchmark}
\jmh{Can you get the figures to color ``eventual'' and ``linearizable'' the same in both?}
  In this benchmark, we measure the scale-out of segmented
  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} while
  we vary the number of servers that we use to replicate our object. When we
  replicate with $n$ servers, we use $3n$ clients (the $3$ colocated clients on
  each server) as part of the workload. The results are shown in
  \figref{VaryNodes}.

### 2018-05-01 doc/paper/sections/interactive_decision_procedure.tex
Couple more small tweaks.
took a pass

\section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\Helper}{\textsf{Helper}}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}
\newcommand{\IsInvConfluent}{\textsc{IsInvConfluent}}

@@

\IsInvConfluent{} assumes access to an \invariantclosure{} decision procedure
\IsIclosed$(O, I)$ which returns a triple (closed, $s_1$, $s_2$). closed is a
boolean indicating whether $O$ is \Iclosed{}. If closed is true, then $s_1$ and
$s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a counterexample
witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
but $\lnot I(s_1 \join s_2)$ (e.g. $s_1$ and $s_2$ from \exampleref{Z2}).  As
we mentioned earlier, we can (and do) implement the \invariantclosure{}
decision procedure using an SMT solver like Z3~\cite{de2008z3}.

@@

\Helper{} recurses only when one of $s_1$ or $s_2$ is unreachable, so $NR$
grows after every recursive invocation of \Helper{}. Similarly, $R$ continues
to grow as \Helper{} randomly explores the set of reachable states. As the user
sees more and more examples of unreachable and reachable states, it becomes
easier and easier for them to recognize patterns that define which states are
reachable and which are not. As a result, it becomes easier for a user to
augment $NR$ and eliminate a large number of unreachable states from the
invariant. Once $NR$ has been sufficiently augmented to the point that $I - NR$
is a subset of the reachable states, \thmref{IclosureEquivalentIconfluence}
guarantees that the algorithm will terminate after one more call to \IsIclosed.

@@

\textbf{A user can label a reachable state as unreachable.}
In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is not
\sTIconfluent. Thus, a user might falsely believe their distributed object to
be \sTIconfluent{} even though it isn't, and eventually one replica of their
distributed object might enter a state that violates the invariant.
%
We can mitigate this in two ways. First, we can aggressively expand $R$. If a
user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the
user of their mistake. Second, \Helper{} returns true when $O$ is \iclosed{(I -
NR)} for some $NR$, so $O$ \emph{is} \sticonfluent{s_0}{T}{I-NR} (even though
it might not be \sTIconfluent{}). Thus, when a user replicates their
distributed object across a set of servers, they can deploy with the invariant
$I - NR$ instead of $I$. If $NR$ is correct and only contains unreachable
states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
$NR$ is incorrect and contains some \sTIreachable{} states, then some of these
states are artificially made unreachable, but the system is still guaranteed to
never produce a state which violates $I$.

>>>>

\section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
\newcommand{\Helper}{\textsc{Helper}}
\newcommand{\IsIclosed}{\textsc{IsIclosed}}
\newcommand{\IsInvConfluent}{\textsc{IsInvConfluent}}

@@

\IsInvConfluent{} assumes access to an \invariantclosure{} decision procedure
\IsIclosed$(O, I)$ which returns a triple (closed, $s_1$, $s_2$). closed is a
boolean indicating whether $O$ is \Iclosed{}. If closed is true, then $s_1$ and
$s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a counterexample
witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from \exampleref{Z2}).  As
we mentioned earlier, we can (and do) implement the \invariantclosure{}
decision procedure using an SMT solver like Z3~\cite{de2008z3}.

@@

\Helper{} recurses only when one of $s_1$ or $s_2$ is unreachable, so $NR$
grows after every recursive invocation of \Helper{}. Similarly, $R$ continues
to grow as \Helper{} randomly explores the set of reachable states. As the user
sees more and more examples of unreachable and reachable states, it often becomes
easier and easier for them to recognize patterns that define which states are
reachable and which are not. As a result, it becomes easier for a user to
augment $NR$ and eliminate a large number of unreachable states from the
invariant. Once $NR$ has been sufficiently augmented to the point that $I - NR$
is a subset of the reachable states, \thmref{IclosureEquivalentIconfluence}
guarantees that the algorithm will terminate after one more call to \IsIclosed.

@@

\textbf{A user can label a reachable state as unreachable.}
In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is not
\sTIconfluent. Thus, a user might falsely believe their distributed object to
be \sTIconfluent{} even though it isn't, and eventually one replica of their
distributed object might enter a state that violates the invariant.
%
We can mitigate this in two ways. First, we can aggressively expand $R$ automatically. If a
user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the
user of their mistake. Second, \Helper{} returns true when $O$ is \iclosed{(I -
NR)} for some $NR$, so $O$ \emph{is} \sticonfluent{s_0}{T}{I-NR} (even though
it might not be \sTIconfluent{}). Thus, when a user replicates their
distributed object across a set of servers, they can deploy with the invariant
$I - NR$ instead of $I$. If $NR$ is correct and only contains unreachable
states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
$NR$ is incorrect and contains some \sTIreachable{} states, then some of these
states are artificially made unreachable, but the system is still guaranteed to
never produce a state which violates $I$.

### 2018-05-01 doc/paper/sections/introduction.tex
took a pass

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like state machine
replication~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search, moraru2013there, vincent2015designing,
oki1988viewstamped, lamport2005generalized, lamport2006fast} makes the
application designer's life very easy. To the application developer, a strongly
consistent system behaves exactly like a single-threaded system running on a
single node, so reasoning about the behavior of the system is
simple~\cite{herlihy1990linearizability}.  Unfortunately, strong consistency is
at odds with performance. The CAP theorem and PACELC theorem tell us that
strongly consistent systems suffer from higher latency at best and
unavailability at worst~\cite{gilbert2002brewer, brewer2012cap,
abadi2012consistency}. On the other hand, weak consistency models like eventual
consistency~\cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In fact, weak consistency
models strip an application developer of one of the earliest and most effective
tools that is used to reason about the execution of programs: application
invariants~\cite{hoare1969axiomatic, balegas2015towards}. Even if every
transaction executing in a weakly consistent system individually maintains an
application invariant, the system as a whole can produce invariant violating
states.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of \invariantconfluence{} as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
\invariantconfluent{} with respect to an invariant, we can replicate it with
the performance benefits of weak consistency and (some of) the correctness
benefits of strong consistency. However, the task of identifying whether or not
an object actually is \invariantconfluent{} remains a challenge. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
\invariantconfluent{} or not, but ideally we would have a general purpose
program that can automatically determine \invariantconfluence{} for us.
\textbf{The first main thrust of this paper is to design a general-purpose
\invariantconfluence{} decision procedure.}

Designing a general-purpose \invariantconfluence{} decision procedure is
difficult\footnote{Actually, it's impossible. Determining whether an object is
\invariantconfluent{} is, in general, undecidable. Still, we can develop a
decision procedure that works well in the common case.}. Many existing
approaches instead develop a decision procedure for \invariantclosure{}, a
sufficient (but not necessary) condition for
\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
approaches check to see if an object is \invariantclosed{}. If it is, then they
conclude that it is also \invariantconfluent{}. If it's not, then the current
approaches are unable to conclude anything about whether or not the object is
\invariantconfluent{}. In this paper, we take a step back and study the
underlying reason \emph{why} \invariantclosure{} is not necessary for
\invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
property about reachability, about whether certain values of our replicated
object can actually be obtained in the execution of a system. We discover that
unreachable, yet invariant satisfying, states cause \invariantclosure{} to be
an unnecessary condition. Using this understanding, we construct a set of
modest conditions under which \invariantclosure{} and \invariantconfluence{}
are in fact \emph{equivalent}, allowing us to reduce the problem of determining
\invariantconfluence{} to that of determining \invariantclosure{}.

Then, we use these conditions to design a general-purpose interactive
\invariantconfluence{} decision procedure. Users iteratively interact with the
decision procedure, eliminating unreachable states from the invariant, moving
towards the conditions under which \invariantclosure{} and
\invariantconfluence{} are equivalent. Meanwhile, the decision-procedure
generates example reachable and unreachable states which help the user
recognize patterns that describe reachability. We also leverage our intuitions
about reachability to develop a new sufficient condition for
\invariantconfluence{}, dubbed \mergereducibility. \Mergereducibility{} covers
some cases that are not covered by \invariantclosure{}, and it can be
automatically checked without any user interaction.

\textbf{The second main thrust of this paper is to take a step beyond
\invariantconfluence{} and develop a generalization of \invariantconfluence{}
called segmented \invariantconfluence{}.} While \invariantconfluence{}
characterizes objects that can be replicated \emph{without any} coordination,
segmented \invariantconfluence{} allows us to replicate
non-\invariantconfluent{} objects with \emph{minimal coordination}. The main
idea is that invariant satisfying states are divided into a number of segments,
and the set of allowable transactions within each segment is restricted.
Within a segment, servers act without any coordination, synchronizing only to
transition across segment boundaries. This design highlights the trade-off
between application complexity and coordination-freedom; more complex
applications require more segments which require more coordination, and
vice-versa.

>>>>

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like distributed transactions~\cite{bernstein1981concurrency,mohan1986transaction} or state machine
replication~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search, moraru2013there, vincent2015designing,
oki1988viewstamped, lamport2005generalized, lamport2006fast} makes the
application designer's life very easy. To the application developer, a strongly
consistent system behaves exactly like a single-threaded system running on a
single node, so reasoning about the behavior of the system is
simple~\cite{herlihy1990linearizability}.  Unfortunately, strong consistency is
at odds with performance. The CAP theorem and PACELC theorem tell us that
strongly consistent systems suffer from higher latency at best and
unavailability at worst~\cite{gilbert2002brewer, brewer2012cap,
abadi2012consistency}. On the other hand, weak consistency models like eventual
consistency~\cite{vogels2009eventually}, PRAM
consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
high availability and low latency, but they put a tremendous burden on the
application designer to reason about the complex interleavings of operations
that are allowed by these weak consistency models. In particular, weak consistency
models strip an application developer of one of the earliest and most effective
tools that is used to reason about the execution of programs: application
invariants~\cite{hoare1969axiomatic, balegas2015towards} and integrity constraints~\cite{IntegrityConstraintCitations}. Even if every
transaction executing in a weakly consistent system individually maintains an
application invariant, the system as a whole can produce invariant-violating
states.

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of \emph{\invariantconfluence{}} as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
\invariantconfluent{} with respect to an invariant, we can replicate it with
the performance benefits of weak consistency and (some of) the correctness
benefits of strong consistency.

Unfortunately, to date, the task of identifying whether or not
an object actually is \invariantconfluent{} has remained an exercise in human proof generation. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
\invariantconfluent{} or not. Hand-written proofs of this sort are unreasonable to expect from programmers. Ideally we would have a general purpose
program that can automatically determine \invariantconfluence{} for us.
\textbf{The first main thrust of this paper is to design a general-purpose
\invariantconfluence{} decision procedure, and implement it in an interactive system.}

Designing such a general-purpose decision procedure is
difficult\footnote{Actually, it's impossible. Determining whether an object is
\invariantconfluent{} is, in general, undecidable. Still, we can develop a
decision procedure that works well in the common case.}. Many existing
approaches instead develop a decision procedure for \emph{\invariantclosure{}}, a
sufficient (but not necessary) condition for
\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
approaches check to see if an object is \invariantclosed{}. If it is, then they
conclude that it is also \invariantconfluent{}. If it's not, then the current
approaches are unable to conclude anything about whether or not the object is
\invariantconfluent{}. In this paper, we take a step back and study the
underlying reason \emph{why} \invariantclosure{} is not necessary for
\invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
reachability property, checking whether certain values of our replicated
object can actually be obtained in the execution of a system. We discover that
unreachable, yet invariant satisfying, states cause \invariantclosure{} to be
an unnecessary condition. Using this understanding, we construct a set of
modest conditions under which \invariantclosure{} and \invariantconfluence{}
are in fact \emph{equivalent}, allowing us to reduce the problem of determining
\invariantconfluence{} to that of determining \invariantclosure{}.

Then, we use these conditions to design a general-purpose interactive
\invariantconfluence{} decision procedure. Users iteratively interact with the
decision procedure, eliminating unreachable states from the invariant, moving
towards the conditions under which \invariantclosure{} and
\invariantconfluence{} are equivalent. Meanwhile, the decision procedure
generates examples of reachable and unreachable states which help the user
recognize patterns that describe reachability. We also leverage our intuitions
about reachability to develop a new sufficient condition for
\invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{} covers
some cases that are not covered by \invariantclosure{}, and it can be
automatically checked without any user interaction.

\textbf{The second main thrust of this paper is to take a step beyond
\invariantconfluence{} and develop a generalization of \invariantconfluence{}
called segmented \invariantconfluence{}.} While \invariantconfluence{}
characterizes objects that can be replicated \emph{without any} coordination,
segmented \invariantconfluence{} allows us to replicate
non-\invariantconfluent{} objects with only \emph{occasional coordination}. The main
idea is that invariant satisfying states are divided into a number of segments,
and the set of allowable transactions within each segment is restricted.
Within a segment, servers act without any coordination, synchronizing only to
transition across segment boundaries. This design highlights the trade-off
between application complexity and coordination-freedom; more complex
applications require more segments which require more coordination, and
vice-versa.

### 2018-05-01 doc/paper/sections/invariant_closure.tex
Couple more tweaks.
took a pass

\thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
sufficient for \invariantconfluence{}. Intuitively, our system model ensures
that transaction execution preserves the invariant, so if merging states also
preserves the invariant and if our start state satisfies the invariant, then
inductively it is impossible for us to reach a state that doesn't satisfy the
invariant.

@@

  The invariant and the set of reachable states are illustrated in \figref{Z2}
  in which we draw each state $(x, y)$ as a point in space. The invariant
  consists of the second and fourth quadrant, and the reachable states consist
  of the fourth quadrant. From this, it's immediate that the reachable states
  are a subset of the invariant, so $O$ is \invariantconfluent{}. However,
  letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
  \invariantclosed{}. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
  (1, 1)$, we see $\lnot I(s_3)$.
\end{example}

In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
\invariantclosed{}, but $s_1$ is not reachable. This is not particular to
\exampleref{Z2}. In fact, it is fundamentally the reason why
\invariantclosure{} is not equivalent to \invariantconfluence{}.
\Invariantconfluence{} is, at its core, a property of reachable states, but
\invariantclosure{} is completely ignorant of reachability. As a result,
invariant satisfying yet unreachable states like $s_1$ are the key hurdle
preventing \invariantclosure{} from being equivalent to \invariantconfluence{}.
This is formalized by \thmref{IclosureEquivalentIconfluence}

\begin{theorem}\thmlabel{IclosureEquivalentIconfluence}
  Consider an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$. If the invariant is a subset of the
  reachable states (i.e. $I \subseteq \setst{s \in
  S}{\sTIreachablepredicate{s}}$), then
  \[
    \text{$I(s_0)$ and $O$ is \Iclosed{}} \iff \text{$O$ is \sTIconfluent{}}
  \]
\end{theorem}
\begin{proof}
  The forward direction of \thmref{IclosureEquivalentIconfluence} follows
  immediately from \thmref{IclosureImpliesIconfluence}. The backward direction
  holds because any two invariant satisfying states $s_1$ and $s_2$ must be
  reachable (by assumption), so their join $s_1 \join s_2$ is also reachable.
  And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
  \join s_2$, satisfy the invariant.
\end{proof}

>>>>

\thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
sufficient for \invariantconfluence{}. Intuitively, recall that our system model ensures
that transaction execution preserves the invariant, so if merging states also
preserves the invariant and if our start state satisfies the invariant, then
inductively it is impossible for us to reach a state that doesn't satisfy the
invariant.

@@

  The invariant and the set of reachable states are illustrated in \figref{Z2}
  in which we draw each state $(x, y)$ as a point in space. The invariant
  consists of the second and fourth quadrant, while the reachable states consist only
  of the fourth quadrant. From this, it is immediate that the reachable states
  are a subset of the invariant, so $O$ is \invariantconfluent{}. However,
  letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
  \invariantclosed{}. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
  (1, 1)$, we see $\lnot I(s_3)$.
\end{example}

In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
\invariantclosed{}, but $s_1$ is not reachable. This is not particular to
\exampleref{Z2}. In fact, it is fundamentally the reason why
\invariantclosure{} is not equivalent to \invariantconfluence{}.
\Invariantconfluence{} is, at its core, a property of reachable states, but
\invariantclosure{} is completely ignorant of reachability. As a result,
invariant-satisfying yet unreachable states like $s_1$ are the key hurdle
preventing \invariantclosure{} from being equivalent to \invariantconfluence{}.
This is formalized by \thmref{IclosureEquivalentIconfluence}.

\begin{theorem}\thmlabel{IclosureEquivalentIconfluence}
  Consider an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$. If the invariant is a subset of the
  reachable states (i.e. $I \subseteq \setst{s \in
  S}{\sTIreachablepredicate{s}}$), then
  \[
    (\text{$I(s_0)$ and $O$ is \Iclosed{}}) \iff \text{$O$ is \sTIconfluent{}}
  \]
\end{theorem}
\begin{proof}
  The forward direction of \thmref{IclosureEquivalentIconfluence} follows
  immediately from \thmref{IclosureImpliesIconfluence}. The backward direction
  holds because any two invariant satisfying states $s_1$ and $s_2$ must be
  reachable (by assumption), so their join $s_1 \join s_2$ is also reachable.
  And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
  \join s_2$, satisfy the invariant.
\end{proof}

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
took a pass

\begin{example}\examplelabel{CounreachableExample}
  Let $O = (\ints^3 \times \ints^3, \join)$ be a
  PN-Counter~\cite{shapiro2011comprehensive} replicated on three machines.
  Every state $s = $ $(p_1, p_2, p_3),$ $(n_1, n_2, n_3)$ represents the
  integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment and decrement
  the counter, the $i$th server increments $p_i$ and $n_i$ respectively, and
  $\join$ computes an element-wise maximum. Our start state $s_0 = (0, 0, 0),
  (0, 0, 0)$; our set $T$ of transactions consists of increment and decrement;
  and our invariant $I$ is that the value of $s$ is non-negative.

  Applying \algoref{InteractiveDecisionProcedure}, \IsIclosed{} returns false
  with the states $s_1 = (1, 0, 0), (0, 1, 0)$ and $s_2 = (1, 0, 0), (0, 0,
  1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
  \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
  concurrent decrements, which we can forbid in a simple one-segment
  segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
  \IsIclosed{} again returns false with same states $s_1$ and $s_2$. This time,
  however, the user recognizes that the two states are not
  \ticoreachable{T^+}{I}. The user refines $NR$ with the observation that two
  states are coreachable if and only if they have the same values of $n_1, n_2,
  n_3$. After this, \IsIclosed{} and
  \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
\end{example}

\subsection{Discussion}
There are a couple of things worth noting about segmented
\invariantconfluence{}, its system model, and its decision procedure.
%
First, \invariantconfluence{} is a very coarse-grained property. If an object
is \invariantconfluent{}, then we can replicate it with no coordination. If it
is not \invariantconfluent{}, then we have no guarantees. There's no
in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
more fine-grained property that can be applied to applications with varying
degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
complex applications that require a large number of segments and to simple
applications with a smaller number of segments, whereas \invariantconfluence{}
only provides guarantees to applications that can be segmented into a single
segment.

@@

Third, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination it requires. First, if a server with
state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
to execute, and $t(s_i)$ violates the active invariant but not the global
invariant, instead of initiating a round of global coordination, $p_i$ can
simply buffer $t$ for re-execution at a later time. While this increases the
latency required to execute $t$, it's possible that after other transactions
are executed, re-executing $t$ may lead to a state that either satisfies the
active invariant or violates the global invariant. In either case, a round of
global coordination is avoided. Similarly, servers can buffer transactions that
require global coordination, executing an entire batch of these transactions
during a single round of global coordination.

@@

Fifth, while a segmented \invariantconfluence{} decision procedure can help
decide whether or not an object is segmented \invariantconfluent{}, it cannot
currently help construct the segmentation. In the future, we plan to explore
ways by which we can automatically suggest segmentations to the application
designer.

>>>>

\begin{example}\examplelabel{CounreachableExample}
  Let $O = (\ints^3 \times \ints^3, \join)$ be an object that separately keeps positive and negative integer counts (dubbed a
  PN-Counter~\cite{shapiro2011comprehensive}), replicated on three machines.
  Every state $s = $ $(p_1, p_2, p_3),$ $(n_1, n_2, n_3)$ represents the
  integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment or decrement
  the counter, the $i$th server increments $p_i$ or $n_i$ respectively, and
  $\join$ computes an element-wise maximum. Our start state $s_0 = (0, 0, 0),
  (0, 0, 0)$; our set $T$ of transactions consists of increment and decrement;
  and our invariant $I$ is that the value of $s$ is non-negative.

  Applying \algoref{InteractiveDecisionProcedure}, \IsIclosed{} returns false
  with the states $s_1 = (1, 0, 0), (0, 1, 0)$ and $s_2 = (1, 0, 0), (0, 0,
  1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
  \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
  concurrent decrements, which we can forbid in a simple one-segment
  segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
  \IsIclosed{} again returns false with same states $s_1$ and $s_2$. This time,
  however, the user recognizes that the two states are not
  \ticoreachable{T^+}{I} \jmh{Clarify here how the user recognizes this ... I believe that you're saying that the definition of $\Sigma$ implies that decrement requires coordination hence $n_i$ are kept consistent hence disagreeing $n_i$ isn't possible.} The user refines $NR$ with the observation that two
  states are coreachable if and only if they have the same values of $n_1, n_2,
  n_3$. After this, \IsIclosed{} and
  \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
\end{example}

\subsection{Discussion}
There are a few things worth noting about segmented
\invariantconfluence{}, its system model, and its decision procedure.
%
First, \invariantconfluence{} is a very coarse-grained property. If an object
is \invariantconfluent{}, then we can replicate it with no coordination. If it
is not \invariantconfluent{}, then we have no guarantees. There's no
in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
more fine-grained property that can be applied to applications with varying
degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
complex applications that require a large number of segments and to simple
applications with a smaller number of segments, whereas \invariantconfluence{}
only provides guarantees to applications that can be segmented into a single
segment.

@@

Third, we can integrate a couple of optimizations into our system model to
further reduce the amount of coordination it requires. To begin, if a server with
state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
to execute, and $t(s_i)$ violates the active invariant but not the global
invariant, instead of initiating a round of global coordination, $p_i$ can
simply buffer $t$ for re-execution at a later time. While this increases the
latency required to execute $t$, it's possible that after other transactions
are executed, re-executing $t$ may lead to a state that either satisfies the
active invariant or violates the global invariant. In either case, a round of
global coordination is avoided. Similarly, servers can buffer transactions that
require global coordination, executing an entire batch of these transactions
during a single round of global coordination.

@@

Fifth, while a segmented \invariantconfluence{} decision procedure can help
decide whether or not an object is segmented \invariantconfluent{}, it cannot
currently help construct a segmentation. In the future, we plan to explore
ways by which we can automatically suggest segmentations to the application
designer.
